---
title: "Writing Checks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing Checks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(chex)
```

## Basic checks

At its core, checks are just predicate functions, which return `TRUE` or `FALSE`
values, which are converted to pass and fail results. rlang-style lambdas are
also accepted. These checks are passed to `check_that()` along with an input
value.

```{r}
is_count <- function(x) {
  is.integer(x) & x > 0
}

check_that(1L, is_count, ~.x == 2L)
```

Checks can also return multiple logical values, which are converted into 
multiple results.

```{r}

check_that(-3:3, is_count)
```


## Descriptions

In the examples above, the check results show the description (e.g. `is_count`)
in addition to the pass or fail value. This description can be customized a few
different ways.

### Named checks

The easiest way to add a custom description is to name your check when calling
`check_that()`:

```{r}
x <- 1L
check_that(x, "`x` is a count" = is_count)
```


### set_description()

If you want to add a more permanent custom description, use the 
`set_description()` function.

```{r}
is_count_set <- set_description(is_count, "`x` is a count")
description(is_count_set)

check_that(x, is_count_set)
```

When you name a check which already has a description set by `set_description()`
the name will take precedence.

```{r}
check_that(x, "overriding set description" = is_count_set)
```


### Dynamic descriptions

Sometimes in advanced use-cases, the description needs to be set dynamically 
from within the check function itself. This can be accomplished using, the
`current_check()` function along with `set_description()`, which we saw above:

```{r}
is_count_dynamic <- function(x) {
  set_description(current_check(), "dynamically set")
  is_count(x)
}

check_that(x, is_count_dynamic)
```

When used with a named description, the named description is used:

```{r}
check_that(x, "named descriptions win" = is_count_dynamic)
```


### Returning results

You can also return a result object with a description already attached to it.
This takes higher precedence than any description that was set using
`set_description()` whether beforehand or dynamically, but has lower precedence
than named descriptions.

```{r}
is_count_result <- function(x) {
  set_description(current_check(), "not used")
  result(is_count(x), "x is a count")
}

check_that(x, is_count_result)

check_that(x, "named descriptions always win" = is_count_result)
```


### Custom S3 class

The final method of adding custom descriptions is the most advanced. This 
requires creating a custom S3 class and some S3 methods for the class. For
example: 

```{r}
my_check <- structure(class = c("my_check", "function"), function(x) {
  TRUE 
})
class(my_check)
```

Then you can create a `description()` method for your class:

```{r}
description.my_check <- function(x, ...) {
  "my check custom description"
}
description(my_check)

check_that(mtcars, my_check)
```

This will work, unless your custom S3 class also returns a result object. In 
the examples above we see that returning a result takes precedence over a set
description. However, with a custom S3 class, you can also write a `run_check()`
method to further customize the behavior:

```{r}
my_check <- structure(class = c("my_check", "function"), function(x) {
  result(TRUE, "this will be used, not the `my check custom description`")
})

# the result description is used, not the check description
check_that(mtcars, my_check) 


# Now define a run_check method. The `check` parameter is your custom check.
run_check.my_check <- function(check, x) {
  result <- check(x) # run your check manually
  set_description(result, "now this overwrites the result description")
}

# now we have full control
check_that(mtcars, my_check)
```

## Reason

Sometimes when a check fails, there will be text beneath the result output which
explains why the check failed. This can happen, for example, when a check throws
an error. This is called the "reason". 

```{r}
check_that(mtcars, function(x) stop("bad. bad. really, really, bad."))
```


You can add a custom reason to your checks, by either throwing and error using
`stop()` or by returning the result of the `FAIL` function:

```{r}
less_than_five <- function(x) {
  if (x >= 5) {
    return(FAIL("x is not less than 5"))
  } 
  TRUE
}

check_that(5, less_than_five)
```


## Check factories

One useful pattern for creating checks is to use a factory. This way you can use
a generic template to create many parameterized check functions. For example,
the `less_than_five()` check in the example above could be parameterized instead
of the `5` being hard-coded:

```{r}
is_less_than_n <- function(n) {
  check <- function(x) {
    if (x >= n) {
      return(FAIL(paste0("x is not be less than ", n, ": x = ", x)))
    }
    TRUE
  }
  set_description(check, paste("x is less than", n))
}

check_that(7, is_less_than_n(n = 5), is_less_than_n(n = 10))
```

